Sound and Automated Synthesis of Digital Stabilizing Controllers for Continuous Plants \thanks{Supported by EPSRC grant EP/J012564/1, ERC project 280053 (CPROVER) and the H2020 FET OPEN SC$^2$.}

### 2017-02-11 hscc2017/paper.tex
hscc17

CEGIS-based control synthesis requires a formal verifier to check whether a
candidate controller meets the requirements when combined with the plant. 
We use the Digital-System Verifier (DSVerifier)~\cite{IsmailBCFF15} for this
purpose.  It checks the stability of closed-loop control systems and
considers finite-word length (FWL) effects in the digital controller, and
uncertainty parameters in the plant model (plant intervals)~\cite{Bessa16}. 
Let $C(z)$ be a digital controller and $G(z)$ be a digital representation of the plant, 
given as
%
\begin{align}
\small
\label{controller_plant_tf}
C(z)&=\frac{C_n(z)}{C_d(z)}=\frac{\beta_{0}+\beta_{1}z^{-1}+...+\beta_{M_C}z^{-M_C}}{\alpha_{0}+\alpha_{1}z^{-1}+...+\alpha_{N_C}z^{-N_C}}, \\
%\label{plant_tf}
G(z)&=\frac{G_n(z)}{G_d(z)}=\frac{b_{0}+b_{1}z^{-1}+...+b_{M_G}z^{-M_G}}{a_{0}+a_{1}z^{-1}+...+a_{N_G}z^{-N_G}}.
\end{align}
%
where $\vec{\beta}$ and $\vec{\alpha}$ are vectors containing the
controller's coefficients; similarly, $\vec{b}$ and $\vec{a}$ denote the
plant's coefficients; 
and finally $N_{(\cdot)}$ and $M_{(\cdot)}$ indicate the order of the polynomials,  
and we require in particular that $N_G \geq M_G$.

@@

Next, we describe the specific property that we pass to the program
synthesizer as the specification $\sigma$.  There are two different
algorithms in DSVerifier that can be used for stability
verification~\cite{daes20161, Bessa16}, one based on Schur's decomposition
and another one based on Jury's criteria~\cite{astrom1997computer}.  Here,
we choose Jury's method~\cite{astrom1997computer}, due to its efficiency,
and we use this method to check the stability in the $z$-domain for the
characteristic polynomial $S(z)$ defined in~\eqref{eq:internal_stab_lemma}.
%
% of the matrix $\left( \begin{array}{c} Y \\ U \end{array}\right)$, where $Y(z)$ is the closed-loop system output and U(z) is the controller output. 
%% \red{[unclear: what matrix? We need to clarify the relationship
%% between the char. poly. $S(z)$ and the later quantities $\Delta
%% N_{G}(z), \Delta D_{G}(z)$ and corresponding controller's
%% structure.]}.
%
We~consider the following form for $S(z)$:
%
$$
S(z) = a_0z^N+a_1z^{N-1}+\cdots+a_{N-1}z+a_N=0, a_0\neq0
$$
%
%\cdsay{is this for the matrix $\left( \begin{array}{c} Y \\ U \end{array}\right)$?}

% Assuming that $\hat{G}(z) = \frac{\hat{N_G}(z)}{\hat{D_G}(z)}$
% and $\hat{C}(z) = \frac{\hat{N_C}(z)}{\hat{D_C}(z)}$, then:
% $$S(z) = FWL[N_C(z)] \times \hat{N_G}(z) + FWL[D_C(z)] \times \hat{D_G}(z)$$

Next, the following Jury matrix
$M = [m_{ij}]_{(2N-2)\times N}$ is built from $S(z)$ coefficients:
%
$$
M=\left( 
\begin{array}{c}
V^{(0)}\\
V^{(1)}\\
\vdots\\
V^{(N-2)}
\end{array}
\right), 
$$
%
where $V^{(k)} = [v^{(k)}_{ij} ]_{2\times N}$ such that:
%
$$
v_{ij}^{(0)}=\left\{
\begin{array}{ll}
a_{j-1}, & \texttt{if}~i=1\\
v_{(1)(N-j+1)}^{0},&\texttt{if}~i=2
\end{array}
\right.
$$
%
$$
v_{ij}^{(k)}=\left\{
\begin{array}{ll}
0,&\texttt{if}~j>n-k\\
v_{1j}^{(k-1)}-v_{2j}^{(k-1)} . \frac{v_{11}^{(k-1)}}{v_{21}^{(k-1)}}, & \texttt{if}~j\leq n-k ~\texttt{and}~i=1\\
v_{(1)(N-j+1)}^{k},& \texttt{if}~j\leq n-k ~\texttt{and}~i=2\\
\end{array}
\right.
$$
%
where $k \in \mathbb{Z}$, such that $0 < k < N - 2$. 
$S(z)$ is the
characteristic polynomial of a stable system if and only if the following four conditions hold:
%\begin{itemize}
%\item 
$R_1: S(1) > 0$;
%\item 
$R_2: (−1)^N S(−1) > 0$;
%\item 
$R_3: |a_0| < a_N$;
%\item 
$R_4: m_{11} > 0 \wedge\allowbreak
      m_{31}>0 \wedge\allowbreak
      m_{51}>0 \wedge \ldots \wedge\allowbreak
      m_{(2N{-}3)(1)}>0$.
%\end{itemize}
%
The stability property is then encoded by a constraint of the form:
$
\phi_\mathit{stability} \equiv (R_1 \wedge R_2 \wedge R_3 \wedge R_4).
$
%\cdsay{figure out what is existentially quantified for the final $\sigma$.} 
%% \red{[Yes, intuitively all clear, however for the sake of clarity I would like to see the explicit mapping between the above formula and the previously used specification 
%% $\sigma(x, P)$, with clear elucidation of the relationship between the domain of $z$ and that of $x$.]}

@@

Since the controller is synthesized in the $\mathbb{R}\langle I,F \rangle$
domain,
%
$$\forall i \leq N_C, j \leq M_C\ \  \tilde \beta_{i} \equiv \beta_{i} \wedge \tilde \alpha_{j} \equiv \alpha_{j} \Leftrightarrow \tilde C(z) \equiv C(z)$$
%
However, given a real plant $\hat{G}(z)$, and a discretizing function
%
\begin{align}
\small
\label{FWL_tf}
\mathcal{F}&\langle I_p,F_p \rangle : \mathbb{R} \rightarrow  \mathbb{R}\langle I_p,F_p \rangle\\
&\triangleq \tilde x = \mathcal{F}\langle I_p,F_p \rangle(x) : \tilde x=x-(x\bmod\tilde x)\nonumber\\
\mathcal{F}_n&\langle I_p,F_p \rangle(P \in \mathcal{P}) \\
&=\tilde P \in \mathcal{P}^{n}\langle I_p,F_p \rangle : c_i \in \vec{P} \wedge \tilde c_i \in \vec{\tilde{P}}=\mathcal{F}\langle I_p,F_p \rangle(c_i)  \nonumber
\end{align}
%
we calculate $\tilde G(z)=\mathcal{F}_{M_G,N_G}\langle I_p,F_p \rangle(\hat{G}(z))$
%
\begin{align}
% \label{digital_controller_tf}
% \tilde C(z)&=\frac{(\hat{\beta}_{0}+\Delta_q \hat{\beta}_{0}) +...+(\hat{\beta}_{M_C}+\Delta_b \hat{\beta}_{M_C})z^{-M_C}}{(\hat{\alpha}_{0}+\Delta_q \hat{\alpha}_{0})+...+(\hat{\alpha}_{N_C}+\Delta_q \hat{\alpha}_{N_C})z^{-N_G}} \nonumber \\
% &= C(z) =\hat{C}(z) + \Delta{C}_b(z)\\
%
\label{digital_plant_tf}
\tilde G(z)&=\frac{(\hat{b}_{0}+\Delta_b \hat{b}_{0}) +...+(\hat{b}_{M_G}+\Delta_b \hat{b}_{M_G})z^{-M_G}}{(\hat{a}_{0}+\Delta_b \hat{a}_{0})+...+(\hat{a}_{N_G}+\Delta_b \hat{a}_{N_G})z^{-N_G}} \nonumber \\
&\Rightarrow \vec{\tilde G} =\vec{\hat{G}}+\Delta_b{\vec{G}}=\vec{G}+\Delta_p{\vec{G}}+\Delta_b{\vec{G}}.
\end{align}
%
where $\Delta_b\hat{c}_i=\hat{c}_i\bmod \tilde{c}_i$
and $\Delta_b{G}$ represent the plant uncertainty caused by
the rounding off effect.  We encompass all the uncertainty as
$\Delta{\vec{G}}=\Delta_p{\vec{G}}+\Delta_b{\vec{G}}$. 
%
As the cost of SAT solving increases with in the size of the problem
instance, our algorithm tries to solve the problem first for small $I_p$,
$F_p$, iteratively increasing the precision if it is insufficient.

\paragraph{Our synthesis problem}
The synthesis problem we are trying to solve is the following:
find a digital controller $C(z)$ %denoted by  $N_c$ and $D_c$ 
that makes the closed-loop system stable 
for all possible uncertainties 
%$\Delta{\vec{p}}$ (\ref{eq:uncertainty}).
$\tilde G(z)$ (\ref{digital_plant_tf}).
%, where we precisely describe 
%$(\Delta N_G, \Delta D_G)$. 
% Note that, similar to the rest of the paper, we use the notation 
% $\tilda{N_G}(z) = N_G(z) + \Delta N_G(z)$ and 
% $\hat{N_C}(z) = N_C(z) + \Delta N_C(z)$.
When mapping back to the notation used for describing the general architecture 
of the program synthesizer, the controller $C(z)$ denotes $P$ and 
$\tilde G(z)$ represents $x$.

@@

The {\sc synthesize} phase uses BMC to 
compute a solution $C(z)$. % An important remark is that this 
% controller is synthesised in the $\mathbb{R}\langle I,F \rangle$
% domain.
There are two approaches for the {\sc verify} phase.
% based on the described model.  
The first approach uses interval arithmetic \cite{moore1966interval}
to represent the coefficients
$[{c}_i-\Delta_p{c}_i-\Delta_b{c}_i,\,\allowbreak
{c}_i+\Delta_p{c}_i-\Delta_b{c}_i+(2^{-F_p})]$ 
and rounds operations outwards during calculations.  This effectively allows
us to simultaneously evaluate the whole collection of plants $\mathfrak{G}$
plus the effects of numeric calculations in the engine.  Synthesized
controllers using this verifier will always be stable for all plants in the
family.
% and only generate counterexamples when the precision of the
%engine is insufficient (\emph{ie} no stable plant could be found). 
Our preliminary experiments showed that a synthesis approach
using this verification engine has poor performance and we
%Given the high computational cost of this
%approach, 
therefore designed a second approach, which we describe below.
Our experimental results described in Section~\ref{sec:experiments}
%try both approaches and provide the time for the fastest. 
show that the speedup yielded by the second approach 
is in most cases of at least two orders of magnitude.

>>>>

CEGIS-based control synthesis requires a formal verifier to check whether a
candidate controller meets the requirements when combined with the plant. 
We use the Digital-System Verifier (DSVerifier)~\cite{IsmailBCFF15} \red{[unclear difference btw dssynth and dsverifier]} for this
purpose.  It checks the stability of closed-loop control systems and
considers finite-word length (FWL) effects in the digital controller, and
uncertainty parameters in the plant model (plant intervals)~\cite{Bessa16}. 
Let $C(z)$ be a digital controller and $G(z)$ be a digital representation of the plant, 
given as
%
\begin{align}
\small
\label{controller_plant_tf}
C(z)&=\frac{C_n(z)}{C_d(z)}=\frac{\beta_{0}+\beta_{1}z^{-1}+...+\beta_{M_C}z^{-M_C}}{\alpha_{0}+\alpha_{1}z^{-1}+...+\alpha_{N_C}z^{-N_C}}, \\
%\label{plant_tf}
G(z)&=\frac{G_n(z)}{G_d(z)}=\frac{b_{0}+b_{1}z^{-1}+...+b_{M_G}z^{-M_G}}{a_{0}+a_{1}z^{-1}+...+a_{N_G}z^{-N_G}}.
\end{align}
%
where $\vec{\beta}$ and $\vec{\alpha}$ are vectors containing the
controller's coefficients; similarly, $\vec{b}$ and $\vec{a}$ denote the
plant's coefficients; 
and finally $N_{(\cdot)}$ and $M_{(\cdot)}$ indicate the order of the polynomials,  
and we require in particular that $N_G \geq M_G$.

@@

Next, we describe the specific property that we pass to the program
synthesizer as the specification $\sigma$.  There are two different
algorithms in \red{DSVerifier} that can be used for stability
verification~\cite{daes20161, Bessa16}, one based on Schur's decomposition
and another one based on Jury's criterion~\cite{astrom1997computer}.  Here,
we choose Jury's method~\cite{astrom1997computer}, due to its efficiency,
and we use this method to check the stability in the $z$-domain for the
characteristic polynomial $S(z)$ defined in~\eqref{eq:internal_stab_lemma}.
%
% of the matrix $\left( \begin{array}{c} Y \\ U \end{array}\right)$, where $Y(z)$ is the closed-loop system output and U(z) is the controller output. 
%% \red{[unclear: what matrix? We need to clarify the relationship
%% between the char. poly. $S(z)$ and the later quantities $\Delta
%% N_{G}(z), \Delta D_{G}(z)$ and corresponding controller's
%% structure.]}.
%
We~consider the following form for $S(z)$:
%
\begin{equation*}
S(z) = a_0z^N+a_1z^{N-1}+\cdots+a_{N-1}z+a_N=0, a_0\neq0. 
\end{equation*}
%\cdsay{is this for the matrix $\left( \begin{array}{c} Y \\ U \end{array}\right)$?}

% Assuming that $\hat{G}(z) = \frac{\hat{N_G}(z)}{\hat{D_G}(z)}$
% and $\hat{C}(z) = \frac{\hat{N_C}(z)}{\hat{D_C}(z)}$, then:
% $$S(z) = FWL[N_C(z)] \times \hat{N_G}(z) + FWL[D_C(z)] \times \hat{D_G}(z)$$
Next, the following Jury matrix
$M = [m_{ij}]_{(2N-2)\times N}$ is built from $S(z)$ coefficients:
%
$$
M=\left( 
\begin{array}{c}
V^{(0)}\\
V^{(1)}\\
\vdots\\
V^{(N-2)}
\end{array}
\right), 
$$
%
where $V^{(k)} = [v^{(k)}_{ij} ]_{2\times N}$ such that:
%
$$
v_{ij}^{(0)}=\left\{
\begin{array}{ll}
a_{j-1}, & \texttt{if}~i=1\\
v_{(1)(N-j+1)}^{0},&\texttt{if}~i=2
\end{array}
\right.
$$
%
$$
v_{ij}^{(k)}=\left\{
\begin{array}{ll}
0,&\texttt{if}~j>n-k\\
v_{1j}^{(k-1)}-v_{2j}^{(k-1)} . \frac{v_{11}^{(k-1)}}{v_{21}^{(k-1)}}, & \texttt{if}~j\leq n-k ~\texttt{and}~i=1\\
v_{(1)(N-j+1)}^{k},& \texttt{if}~j\leq n-k ~\texttt{and}~i=2\\
\end{array}
\right.
$$
%
where $k \in \mathbb{Z}$, such that $0 < k < N - 2$. 
$S(z)$ is the
characteristic polynomial of a stable system if and only if the following four conditions hold:
%\begin{itemize}
%\item 
$R_1: S(1) > 0$;
%\item 
$R_2: (−1)^N S(−1) > 0$;
%\item 
$R_3: |a_0| < a_N$;
%\item 
$R_4: m_{11} > 0 \wedge\allowbreak
      m_{31}>0 \wedge\allowbreak
      m_{51}>0 \wedge \ldots \wedge\allowbreak
      m_{(2N{-}3)(1)}>0$.
%\end{itemize}
%
The stability property is then encoded by a constraint of the form:
$
\phi_\mathit{stability} \equiv (R_1 \wedge R_2 \wedge R_3 \wedge R_4).
$
%\cdsay{figure out what is existentially quantified for the final $\sigma$.} 
%% \red{[Yes, intuitively all clear, however for the sake of clarity I would like to see the explicit mapping between the above formula and the previously used specification 
%% $\sigma(x, P)$, with clear elucidation of the relationship between the domain of $z$ and that of $x$.]}

@@

Since the controller is synthesized in the $\mathbb{R}\langle I,F \rangle$
domain,
%
$$
\forall i \leq N_C, j \leq M_C\ \  \tilde \beta_{i} \equiv \beta_{i} \wedge \tilde \alpha_{j} \equiv \alpha_{j}, \text{ that is } \tilde C(z) \equiv C(z). 
$$
%
However, given a real plant $\hat{G}(z)$, and a discretizing function
%
\begin{align}
\small
\label{FWL_tf}
\mathcal{F}&\langle I_p,F_p \rangle : \mathbb{R} \rightarrow  \mathbb{R}\langle I_p,F_p \rangle\\
&\triangleq \tilde x = \mathcal{F}\langle I_p,F_p \rangle(x) : \tilde x=x-(x\bmod\tilde x)\nonumber\\
\mathcal{F}_n&\langle I_p,F_p \rangle(P \in \mathcal{P}) \\
&=\tilde P \in \mathcal{P}^{n}\langle I_p,F_p \rangle : c_i \in \vec{P} \wedge \tilde c_i \in \vec{\tilde{P}}=\mathcal{F}\langle I_p,F_p \rangle(c_i)  \nonumber
\end{align}
%
we calculate $\tilde G(z)=\mathcal{F}_{M_G,N_G}\langle I_p,F_p \rangle(\hat{G}(z))$, where 
%
\begin{align}
% \label{digital_controller_tf}
% \tilde C(z)&=\frac{(\hat{\beta}_{0}+\Delta_q \hat{\beta}_{0}) +...+(\hat{\beta}_{M_C}+\Delta_b \hat{\beta}_{M_C})z^{-M_C}}{(\hat{\alpha}_{0}+\Delta_q \hat{\alpha}_{0})+...+(\hat{\alpha}_{N_C}+\Delta_q \hat{\alpha}_{N_C})z^{-N_G}} \nonumber \\
% &= C(z) =\hat{C}(z) + \Delta{C}_b(z)\\
%
\label{digital_plant_tf}
\tilde G(z)&=\frac{(\hat{b}_{0}+\Delta_b \hat{b}_{0}) +...+(\hat{b}_{M_G}+\Delta_b \hat{b}_{M_G})z^{-M_G}}{(\hat{a}_{0}+\Delta_b \hat{a}_{0})+...+(\hat{a}_{N_G}+\Delta_b \hat{a}_{N_G})z^{-N_G}} \nonumber \\
&\Rightarrow \vec{\tilde G} =\vec{\hat{G}}+\Delta_b{\vec{G}}=\vec{G}+\Delta_p{\vec{G}}+\Delta_b{\vec{G}}, 
\end{align}
%
where $\Delta_b\hat{c}_i=\hat{c}_i\bmod \tilde{c}_i$, 
and $\Delta_b{G}$ represents the plant uncertainty caused by
the rounding off effect.  We encompass all the uncertainty as
$\Delta{\vec{G}}=\Delta_p{\vec{G}}+\Delta_b{\vec{G}}$. 
%
As the cost of SAT solving increases with in the size of the problem
instance, our algorithm tries to solve the problem first for small $I_p$,
$F_p$, iteratively increasing the precision if it is insufficient.

\paragraph{The synthesis problem}
The synthesis problem we are trying to solve is the following:
find a digital controller $C(z)$ %denoted by  $N_c$ and $D_c$ 
that makes the closed-loop system stable 
for all possible uncertainties 
%$\Delta{\vec{p}}$ (\ref{eq:uncertainty}).
$\tilde G(z)$ (\ref{digital_plant_tf}).
%, where we precisely describe 
%$(\Delta N_G, \Delta D_G)$. 
% Note that, similar to the rest of the paper, we use the notation 
% $\tilda{N_G}(z) = N_G(z) + \Delta N_G(z)$ and 
% $\hat{N_C}(z) = N_C(z) + \Delta N_C(z)$.
When mapping back to the notation used for describing the general architecture 
of the program synthesizer, the controller $C(z)$ denotes $P$ and 
$\tilde G(z)$ represents $x$.

@@

The {\sc synthesize} phase uses BMC to 
compute a solution $C(z)$. % An important remark is that this 
% controller is synthesised in the $\mathbb{R}\langle I,F \rangle$
% domain.
There are two approaches for the {\sc verify} phase.
% based on the described model.  
The first approach uses interval arithmetic \cite{moore1966interval}
to represent the coefficients
$[{c}_i-\Delta_p{c}_i-\Delta_b{c}_i,\,\allowbreak
{c}_i+\Delta_p{c}_i-\Delta_b{c}_i+(2^{-F_p})]$ 
and rounds operations outwards during calculations.  This effectively allows
us to simultaneously evaluate the whole collection of plants $\mathfrak{G}$ \red{[weird new notation needed? used elsewhere?]}
plus the effects of numeric calculations in the engine.  Synthesized
controllers using this verifier will always be stable for all plants in the
family.
% and only generate counterexamples when the precision of the
%engine is insufficient (\emph{ie} no stable plant could be found). 
Our preliminary experiments showed that a synthesis approach
using this verification engine has poor performance and we
%Given the high computational cost of this
%approach, 
therefore designed a second approach, which we describe below.
Our experimental results described in Section~\ref{sec:experiments}
%try both approaches and provide the time for the fastest. 
show that the speedup yielded by the second approach 
is in most cases of at least two orders of magnitude.
